Investigating the Applications and Limits of Single Asymmetric Particle Light Scattering using Limited Detection Schemes

### 2022-11-14 Manuscript.tex
starting to clean up manuscript

\usepackage{amssymb, amsthm}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{wrapfig}
\usepackage{geometry}
\usepackage{fancyref}
\usepackage{lineno}
\usepackage{color}
\usepackage{nomecl}

\journal{Journal of Quantitative Spectroscopy $\&$ Radiative Transfer}

@@

\begin{abstract}
\justifying
While optical trapping is a well understood method for force transduction and detection,characterisation of trapped entities poses a two-fold challenge - one experimental concerning the arrangement of light detectors and the other, theoretical involving solving of the inverse light scattering problem. Combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup. We propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme and demonstrate how a Bayesian inference based analysis of the data simulated to mimic light scattering detection signals in such scenarios maybe used for solving the inverse light scattering problem and help characterise trapped entities. To this end, we discuss the application of our method to infer the instantaneous orientations of an asymmetric microsphere dimer being trapped. We argue that this method can be extended for determining any characteristics of the trapped microstructure that influence the light scattering pattern.
\end{abstract}

@@

\nomenclature{$\hat{s}$}{Dimer's instantaneous orientation}
\nomenclature{$\hat{n}_i$}{Reference Orientation i}

@@

\section{Introduction}
\label{sec:Intro}

@@

\section{Methodology}
\label{sec:Method}
\subsection{Brownian Simulation}
\label{sec:2.1}
We use the Brownian OT package developed by Jerome Fung \textit{et al} \cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric dimer within an optical trap. Brownian OT is unique in that it combines MSTM \cite{Mishchenko1996MSTM} and 'Optical Tweezer Toolbox' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary shaped sphere clusters. We simulated the motion of silica dimer trapped in a 5 mW Gaussian beam; the dimer was comprised of two spheres with radii $1\mu m$ and $ 0.5  \mu m$ respectively. We ran our simulation over the span of 10 seconds, with Brownian OT calculating the dimer's orientation and position every $1 \ ms$.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{./Images/Dimer.png}
	\caption{Example dimer in orientation $\hat{s}$}
	\label{fig:dimer}
\end{figure}

With any trapped particle there are two main sources of motion, Brownian and optical forces. The dimer's Brownian motion ($\Delta q^{\textbf{B}}$) is simulated by generating a set of normally-distributed numbers whose covariance is given by the dimer's diffusion tensor $\textbf{D}$. The diffusion tensor is computed based on Farsund's work \cite{Farsund1996}. For computing the optical force Brownian OT uses the T-Matrix method to calculate the scattering coefficients of the outgoing scattered wave. The T-matrix describes the scattering process by relating the scattering wave to the incident wave:

\begin{align}
	\begin{pmatrix}
		p_{mn} \\
		q_{mn}
	\end{pmatrix}
	= \textbf{T}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

MSTM computes the T-matrix of the dimer, and then passes the T-matrix to \textit{ott} to compute the optical forces. The benefit of this approach is that MSTM can accurately compute the T-matrix for any arbitrary collection of spheres once, and then \textit{ott} can calculate the optical forces and torques. The total displacement is the sum of the optical and Brownian contributions:

\begin{align}
	\langle \Delta q_i^{\textbf{B}} \Delta q_j^{\textbf{B}}\rangle = 2D_{ij} \Delta t \\
	\Delta q^{opt} = \frac{\Delta t}{k_BT}\textbf{D} \cdot \textbf{F} \\
	\Delta q^{tot} = \Delta q^{\textbf{B}} + \Delta q^{opt}
\end{align}

where \textbf{F} is the a 6 component vector that describes the optical force and torque components. Brownian OT returns the dimer's motion as a \textit{csv} file containing for each timestep the $[x,y,z]$ position of the dimer's centre of mass relative to the trap focus, and the orientation represented as a quaternion.

\subsection{Orientation estimation from scattering measurements}
\label{sec:2.2}

We represent the instantaneous orientation of the moving dimer in the trap as a three component unit vector,  $\hat{s}$, pointing from the centre of the larger sphere to the centre of the smaller sphere. A plane wave 'probe' laser beam directed perpendicular to the trapping laser, incident on the dimer, generates a scattering pattern dependent on the dimer's instantaneous orientation,  $I(\hat{s},\ \theta)$. This scattering pattern was computed using MSTM for each time step of our Brownian simulation. To represent a simulated 'measurement', three angles ($\theta_1,\ \theta_2,\ \& \theta_3$) were chosen and the calculated  intensity at each angle k recorded as $I(\hat{s},\ \theta_k)$.

Our goal is to estimate the dimer orientation $\hat{s}$ over time, based on the scattering measurements  $I(\hat{s},\ \theta_k)$: the question then is what set of angles ($\theta_1,\ \theta_2,\ \& \theta_3$) give us the most useful information on $\hat{s}$. To initially simplify the problem, rather than trying to get an exact estimation of $\hat{s}$ we discretise the orientation space into 30 possible reference orientations $\hat{n}_i$ that divide up a unit sphere evenly \cite{Rey2006}: our question becomes, how can we optimise the experiment to identify the reference orientation $\hat{n}_i$ closest to the real orientation $\hat{s}$. (Ultimately our method could be extended to an arbitrarily large number of reference orientations but for clarity and practicability we use a relatively small number here; additionally, the choice of a small number of discrete reference or 'candidate' orientations also accounts for the fact that any experiment will necessarily have a limited precision.)   Using MSTM we compute the scattering that would be produced by the dimer if it were in each of the reference orientations, denoted as $I(\hat{n}_i, \ \theta)$ and obtain 'measurements'  $I(\hat{n}_i, \ \theta_k)$ at our chosen set of  angles $\theta_k$. These raw intensities are normalised according to \eqref{eq:scale}:

\begin{align}
	\label{eq:scale}
	y_k(\hat{n}_i,\ \theta_k) = 
	 \frac{\left[I(\hat{n}_i, \ \theta_k)- \langle I(\hat{n}, \ \theta_k) \rangle \right]} 
	{\langle I^2(\hat{n},\ \theta_k) \rangle -\langle I(\hat{n}, \ \theta_k)\rangle^2}
\end{align}

Where the denominator is simply the standard deviation for our values of $I(\hat{n}, \theta_k)$. The dimer's \emph{actual}  instantaneous scattered intensity based on the real instantaneous orientation $\hat{s}$  is similarly scaled using \eqref{eq:scale} to give $y(\hat{s},\ \theta_k)$. See Table's A.1 and A.2 for numerical values of our reference orientation, raw intensities, and scaled signals. In practice $y(\hat{s}\, \theta_k)$ will have an inherent amount of uncertainty in our measurement due to noise from our detector. We compare our instantaneous signal to our reference signal by assuming that the measurement uncertainty for each detection fibre is Gaussian and is centred about $y_k(\hat{s})$ Combining all three detection angles provides us with the conditional probability that, should the dimer be in orientation $\hat{n}_i$, it would produce a signal $y(\hat{s})$:

\begin{align}
	\label{eq:conditional}
	p_i(y(\hat{s})\parallel\hat{n}_i) = \Pi^3_{k=1}
	(2\pi\sigma_k^2)^{-1/2} 
	e^{-(y_{k}(\hat{n}_i)-y_{k}(\hat{s}))^2/2\sigma_k^2}
\end{align}

Where $\sigma_k$ is the average error in the detected signal from our dimer $y_k(\hat{s})$. We left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model. The signal error was simulated in our model by stochastically adjusting our values of $y_k(\hat{s})$.

\begin{eqnarray*}
	\sigma_k = \epsilon\bar{y}(\hat{n}_k) \\
	\Rightarrow y(\hat{s}_k) = y(\hat{s})_k \pm n\sigma_k \\ 
	n \in \mathbb{R}, n\in[0,1]
\end{eqnarray*}

Where $\epsilon$ is the percent error expected from our detection fibres. Equation~\eqref{eq:conditional} assigns a probability to each reference orientation based on how similar the 'reference orientation signal' $y_{k}(\hat{n}_i)$ is to the actual signal $y_{k}(\hat{s})$ generated by the dimer in its real orientation $\hat{s}$. However, if our values of $y_k(\hat{n}_i)$ are similar to one another the probability distribution from equation~\ref{eq:conditional} will therefore assign similar probability for multiple reference orientations. Bayesian inference allows one to update a probability distribution by accounting for information known prior to making a measurement. We apply this by accounting for our distribution of reference orientations and instantaneous signals.

\begin{align}
	\label{eq:bayes}
	p_i(\hat{n}\parallel y(\hat{s}))&= \frac{p_i(y(\hat{s})\parallel\hat{n})p_i(\hat{n})}{p(y)}
\end{align}

Where $p_i(\hat{n})$ and $p(y)$ are our estimations of the prior distributions of particle orientations and instantaneous signals respectively. The former can be estimated by some Boltzmann distribution based on how far apart reference orientations are in physical space, covered in Section~\ref{sec:2.3}. The latter prior $p(y)$ is the probability of our dimer producing a signal $y_k(\hat{s})$ that falls with our distribution of reference signals $y_k(\hat{n})$. This is given by taking the discrete integral over collection of reference orientations.

\begin{align}
	p(y) = \sum\limits_{i=1}^{30} p_i(y(\hat{s})\parallel\hat{n})p_i(\hat{n})
 \end{align}

@@

\subsection{Testing our Model}
\label{sec:2.3}
Using our simulation from Section~\ref{sec:2.1} we calculated the instantaneous signal from our dimer at each time step using MSTM.  We applied \eqref{eq:bayes} - covered in Section~\ref{sec:2.2} - and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation. The priori of reference orientations $p_i(\hat{n})$ is defined at each time step by accounting for the physical distance between our previous estimation $\hat{n}_{i, \ t - \Delta t}$ and each reference orientation $\hat{n}$

\begin{align}
	r_i(t, \hat{n}_i)= \frac{1}{2}((\hat{n}_{i,\ t} \cdot \hat{n}_{t-\Delta t})-1) \\
	p_i(\hat{n})= \frac{e^{\beta (r_i(t,\hat{n}_i)-1)}}
	{\Sigma_{j=1}^{30}e^{\beta(r_j(t, \hat{n}_i)-1)}}
\end{align}

Where $\beta$ is a weighting factor that controls the significance of our priori estimate. The simulation data from Section~\ref{sec:2.1} was used to check if the estimation was correct by comparing the estimate provided by \eqref{eq:bayes} to the reference orientation closest to the instantaneous orientation, denoted as $\hat{n}_{best}$. To measure the error in the model's estimate the distribution's of $p(\hat{n}\parallel y(\hat{s}))$ and an ideal result, $p_ideal$, were compared. An ideal distribution is one that is 0 for every $\hat{n}$ apart from the correct result.

\begin{align}
	p_{ideal} = 
	\begin{cases}
		1 & \text{when $\hat{n}_i$ = $\hat{n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}

The distribution from \eqref{eq:bayes} will assign some non-zero probability to every reference orientation. The model's error was measured by calculating the Kullback-Leibler ($K_l$) divergence between the two distributions

\begin{align}
	K_{l, \#}(p_i(\hat{n}\parallel y_k(\hat{s})) &\parallel p_{ideal}) =\nonumber \\
	&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{n}_{best}\parallel y(\hat{s}))}
	\right]
\end{align}

Where a more negative value of $K_l$ indicates a poorer estimation of the dimer's instantaneous orientation. By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation. To compare between different iterations of our model the we compare our result of $K_l$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:

\begin{align}
	K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
	K_{l, \ worst} &= -\sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/30} \right] \\
	F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
\end{align}

@@

\subsection{Optimization of parameters}
\label{sec:2.4}
The signal error $\epsilon$ and the weighting factor $\beta$ are known quantities whose influence can be measured directly. However, our choice of angles and how they correlate with $F(K_l)$ is not easily understood Evaluating every possible configuration manually is inefficient, therefore we utilised Ultra-nest to apply a Monte Carlo simulation to our model. Ultra-nest is a sophisticated python package for fitting models with complex parameter correlations \cite{Buchner2016Ultranest}. Ultra-nest implements a variation of Monte Carlo integration, nested-sampling, where a likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).

@@

As can be seen from Figure~\ref{fig:epsilon} there is seemingly a sweet spot for the signal error; with a $5\%$ error giving an inaccurate estimation, and with a $25\%$ error giving an alright estimation but with sudden shifts away from the ideal result. The reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k$ is larger than our signal error. As such the model could not distinguish between different reference orientations and was essentially just picking options at random. This minimum difference determined by our choice of angles and sets a lower limit of how accurate our signals can be.  In the future we hope to look at increasing the number of reference orientations to see what the true lower bound is.

@@

As shown by \ref{fig:beta} eventually $\beta$ begins to over correct our estimate to the extent that we are a full radian away from the correct result. The exact value of $\beta$ that will optimise our performance varies for each choice of angle, therefore we can use it as variable for finely tuning our results to suit our choice of angles and our signal error. In the future we want to also investigate the possibility of extending our prior $p(\hat{n})$ so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate if we can investigate the auto correlation of our dimer's motion.

@@

\section{Conclusion}
\label{sec:Conclusion}
We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern. In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity. The MSTM package is flexible in calculating the light scattering for a variety of micro-particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy. This has potential to be applied to characterise particle size and shape that are changing while in the optical trap. The authors would like to acknowledge and thank the support for this research from the funding provided by the Leverhulme Trust.\\

\noindent \textbf{Acknowledgement.} The authors thank the support for this research from the funding provided by the Leverhulme Trust. \\

@@

\bibliography{bib} 
\bibliographystyle{ieeetr}

>>>>

\usepackage{amssymb, amsthm}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{wrapfig}
\usepackage{geometry}
\usepackage{fancyref}
\usepackage{lineno}
\usepackage{color}
\usepackage{nomencl}

\journal{Journal of Quantitative Spectroscopy \& Radiative Transfer}

@@

\begin{abstract}
  % \justifying
  %
  While optical trapping is a well understood method for force
  transduction and detection,characterisation of trapped entities
  poses a two-fold challenge --- one experimental concerning the
  arrangement of light detectors and the other, theoretical involving
  solving of the inverse light scattering problem. Combining static
  light scattering techniques with optical trapping poses significant
  engineering challenges due to the space constraints in a
  conventional optical trapping setup.  We propose here a plausible
  scenario of detecting scattered light from an optically trapped
  asymmetric microstructure using a novel, multi-angle, optical-fibre
  based detection scheme and demonstrate how a Bayesian inference
  based analysis of the data simulated to mimic light scattering
  detection signals in such scenarios maybe used for solving the
  inverse light scattering problem and help characterise trapped
  entities.  To this end, we discuss the application of our method to
  infer the instantaneous orientations of an asymmetric microsphere
  dimer being trapped. We argue that this method can be extended for
  determining any characteristics of the trapped microstructure that
  influence the light scattering pattern.
\end{abstract}

@@

\nomenclature{$\bar{\bf s}$}{Dimer's instantaneous orientation}
\nomenclature{$\hat{\bf n}_i$}{Reference Orientation i}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:Intro}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
\label{sec:Method}

\subsection{Brownian Simulation}
\label{sec:2.1}

We use the Brownian OT package developed by Fung~\textit{et~al}.
\cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric
dimer within an optical trap.  Brownian OT combines MSTM
\cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox''
(\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
shaped sphere clusters.  We simulated the motion of silica dimer
trapped in a $5$\,mW Gaussian beam; the dimer was comprised of two
spheres with radii $1\,{\rm \mu m}$ and $ 0.5\,{\rm \mu m}$,
respectively.  We ran our simulation over the span of 10 seconds, with
Brownian OT calculating the dimer's orientation and position every
$1 \ ms$.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{./Images/Dimer.png}
	\caption{Example dimer in orientation $\bar{\bf s}$}
	\label{fig:dimer}
\end{figure}

With any trapped particle there are two main sources of motion,
Brownian and optical forces.  The dimer's Brownian motion
($\Delta q^{\textbf{B}}$) is simulated by generating a set of
normally-distributed numbers whose covariance is given by the dimer's
diffusion tensor $\textbf{D}$.  The diffusion tensor is computed based
on Farsund's work \cite{Farsund1996}. For computing the optical force
Brownian OT uses the T-Matrix method to calculate the scattering
coefficients of the outgoing scattered wave.  The T-matrix describes
the scattering process by relating the scattering wave to the incident
wave:
\begin{align}
	\begin{pmatrix}
		p_{mn} \\
		q_{mn}
	\end{pmatrix}
	= \textbf{T}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}
MSTM computes the T-matrix of the dimer, and then passes the T-matrix
to \textit{ott} to compute the optical forces. The benefit of this
approach is that MSTM can accurately compute the T-matrix for any
arbitrary collection of spheres once, and then \textit{ott} can
calculate the optical forces and torques. The total displacement is
the sum of the optical and Brownian contributions:
\begin{align}
  \langle \Delta q_i^{\textbf{B}} \Delta q_j^{\textbf{B}}\rangle
  &= 2D_{ij} \Delta t
  \\
  \Delta q^{opt} &= \frac{\Delta t}{k_BT}\textbf{D} \cdot \textbf{F}
  \\
\Delta q^{tot} &= \Delta q^{\textbf{B}} + \Delta q^{opt}
\end{align}
where \textbf{F} is the a 6-component vector that describes the
optical force and torque components.  Brownian OT returns the dimer's
motion as a \textit{csv} file containing for each timestep the
$[x,y,z]$ position of the dimer's centre of mass relative to the trap
focus, and the orientation represented as a quaternion.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Orientation estimation from scattering measurements}
\label{sec:2.2}

We represent the instantaneous orientation of the moving dimer in the
trap as a three component unit vector, $\hat{\bf s}$, pointing from
the centre of the larger sphere to the centre of the smaller sphere.
A plane wave ``probe'' laser beam directed perpendicular to the
trapping laser, incident on the dimer, generates a scattering pattern
dependent on the dimer's instantaneous orientation,
$I(\bar{\bf s},\ \theta)$. This scattering pattern was computed using MSTM
for each time step of our Brownian simulation.  To represent a
simulated 'measurement', three angles ($\theta_1$, $\theta_2$,
$\theta_3$) were chosen and the calculated intensity at each angle k
recorded as $I(\hat{\bf s}, \theta_k)$.

Our goal is to estimate the dimer orientation $\bar{\bf s}$ over time,
based on the scattering measurements $I(\bar{\bf s},\ \theta_k)$: the
question then is what set of angles ($\theta_1$, $\theta_2$,
$\theta_3$) give us the most useful information on $\hat{\bf s}$.  To
initially simplify the problem, rather than trying to get an exact
estimation of $\hat{\bf s}$ we discretise the orientation space into
$30$ possible reference orientations $\hat{\bf n}_i$ that divide up a
unit sphere evenly \cite{Rey2006}: our question becomes, how can we
optimise the experiment to identify the reference orientation
$\hat{\bf n}_i$ closest to the real orientation $\hat{\bf s}$.
(Ultimately our method could be extended to an arbitrarily large
number of reference orientations but for clarity and practicability we
use a relatively small number here; additionally, the choice of a
small number of discrete reference or 'candidate' orientations also
accounts for the fact that any experiment will necessarily have a
limited precision.)  Using MSTM we compute the scattering that would
be produced by the dimer if it were in each of the reference
orientations, denoted as $I(\hat{\bf n}_i, \ \theta)$ and obtain
'measurements' $I(\hat{\bf n}_i, \ \theta_k)$ at our chosen set of angles
$\theta_k$. These raw intensities are normalised according to
\eqref{eq:scale}:

\begin{align}
	\label{eq:scale}
	y_k(\hat{\bf n}_i,\ \theta_k) = 
	 \frac{\left[I(\hat{\bf n}_i, \ \theta_k)- \langle I(\hat{\bf n}, \ \theta_k) \rangle \right]} 
	{\langle I^2(\hat{\bf n},\ \theta_k) \rangle -\langle I(\hat{\bf n}, \ \theta_k)\rangle^2}
\end{align}

Where the denominator is simply the standard deviation for our values of $I(\hat{\bf n}, \theta_k)$. The dimer's \emph{actual}  instantaneous scattered intensity based on the real instantaneous orientation $\bar{\bf s}$  is similarly scaled using \eqref{eq:scale} to give $y(\bar{\bf s},\ \theta_k)$. See Table's A.1 and A.2 for numerical values of our reference orientation, raw intensities, and scaled signals. In practice $y(\bar{\bf s}\, \theta_k)$ will have an inherent amount of uncertainty in our measurement due to noise from our detector. We compare our instantaneous signal to our reference signal by assuming that the measurement uncertainty for each detection fibre is Gaussian and is centred about $y_k(\bar{\bf s})$ Combining all three detection angles provides us with the conditional probability that, should the dimer be in orientation $\hat{\bf n}_i$, it would produce a signal $y(\bar{\bf s})$:

\begin{align}
	\label{eq:conditional}
	p_i(y(\bar{\bf s})\parallel\hat{\bf n}_i) = \Pi^3_{k=1}
	(2\pi\sigma_k^2)^{-1/2} 
	e^{-(y_{k}(\hat{\bf n}_i)-y_{k}(\bar{\bf s}))^2/2\sigma_k^2}
\end{align}

Where $\sigma_k$ is the average error in the detected signal from our dimer $y_k(\bar{\bf s})$. We left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model. The signal error was simulated in our model by stochastically adjusting our values of $y_k(\bar{\bf s})$.

\begin{eqnarray*}
	\sigma_k = \epsilon\bar{y}(\hat{\bf n}_k) \\
	\Rightarrow y(\bar{\bf s}_k) = y(\bar{\bf s})_k \pm n\sigma_k \\ 
	n \in \mathbb{R}, n\in[0,1]
\end{eqnarray*}

Where $\epsilon$ is the percent error expected from our detection fibres. Equation~\eqref{eq:conditional} assigns a probability to each reference orientation based on how similar the 'reference orientation signal' $y_{k}(\hat{\bf n}_i)$ is to the actual signal $y_{k}(\bar{\bf s})$ generated by the dimer in its real orientation $\bar{\bf s}$. However, if our values of $y_k(\hat{\bf n}_i)$ are similar to one another the probability distribution from equation~\ref{eq:conditional} will therefore assign similar probability for multiple reference orientations. Bayesian inference allows one to update a probability distribution by accounting for information known prior to making a measurement. We apply this by accounting for our distribution of reference orientations and instantaneous signals.

\begin{align}
	\label{eq:bayes}
	p_i(\hat{\bf n}\parallel y(\bar{\bf s}))&= \frac{p_i(y(\bar{\bf s})\parallel\hat{\bf n})p_i(\hat{\bf n})}{p(y)}
\end{align}

Where $p_i(\hat{\bf n})$ and $p(y)$ are our estimations of the prior distributions of particle orientations and instantaneous signals respectively. The former can be estimated by some Boltzmann distribution based on how far apart reference orientations are in physical space, covered in Section~\ref{sec:2.3}. The latter prior $p(y)$ is the probability of our dimer producing a signal $y_k(\bar{\bf s})$ that falls with our distribution of reference signals $y_k(\hat{\bf n})$. This is given by taking the discrete integral over collection of reference orientations.

\begin{align}
	p(y) = \sum\limits_{i=1}^{30} p_i(y(\bar{\bf s})\parallel\hat{\bf n})p_i(\hat{\bf n})
 \end{align}

@@

\subsection{Testing our Model}
\label{sec:2.3}
Using our simulation from Section~\ref{sec:2.1} we calculated the instantaneous signal from our dimer at each time step using MSTM.  We applied \eqref{eq:bayes} - covered in Section~\ref{sec:2.2} - and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation. The priori of reference orientations $p_i(\hat{\bf n})$ is defined at each time step by accounting for the physical distance between our previous estimation $\hat{\bf n}_{i, \ t - \Delta t}$ and each reference orientation $\hat{\bf n}$

\begin{align}
	r_i(t, \hat{\bf n}_i)= \frac{1}{2}((\hat{\bf n}_{i,\ t} \cdot \hat{\bf n}_{t-\Delta t})-1) \\
	p_i(\hat{\bf n})= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
	{\Sigma_{j=1}^{30}e^{\beta(r_j(t, \hat{\bf n}_i)-1)}}
\end{align}
where $\beta$ is a weighting factor that controls the significance of
our priori estimate. The simulation data from Section~\ref{sec:2.1}
was used to check if the estimation was correct by comparing the
estimate provided by \eqref{eq:bayes} to the reference orientation
closest to the instantaneous orientation, denoted as
$\hat{\bf n}_{best}$.  To measure the error in the model's estimate the
distribution's of $p(\hat{\bf n}\parallel y(\bar{\bf s}))$ and an ideal
result, $p_ideal$, were compared. An ideal distribution is one that is
0 for every $\hat{\bf n}$ apart from the correct result.
\begin{align}
	p_{ideal} = 
	\begin{cases}
		1 & \text{when $\hat{\bf n}_i$ = $\hat{\bf n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}
The distribution from \eqref{eq:bayes} will assign some non-zero probability to every reference orientation.  The model's error was measured by calculating the Kullback-Leibler ($K_l$) divergence between the two distributions

\begin{align}
	K_{l, \#}(p_i(\hat{\bf n}\parallel y_k(\bar{\bf s})) &\parallel p_{ideal}) =\nonumber \\
	&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}\parallel y(\bar{\bf s}))}
	\right]
\end{align}
where a more negative value of $K_l$ indicates a poorer estimation of
the dimer's instantaneous orientation. By summing the divergence of
each measurement across the entire simulation we get an evaluation of
how well the model performed in estimating the dimer's orientation. To
compare between different iterations of our model the we compare our
result of $K_l$ to a worst case scenario and evaluate how much it
improves upon this, denoted as $F(K_l)$:
\begin{align}
	K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
	K_{l, \ worst} &= -\sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/30} \right] \\
	F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Optimization of parameters}
\label{sec:2.4}
The signal error $\epsilon$ and the weighting factor $\beta$ are known quantities whose influence can be measured directly. However, our choice of angles and how they correlate with $F(K_l)$ is not easily understood Evaluating every possible configuration manually is inefficient, therefore we utilised Ultra-nest to apply a Monte Carlo simulation to our model. Ultra-nest is a sophisticated python package for fitting models with complex parameter correlations \cite{Buchner2016Ultranest}. Ultra-nest implements a variation of Monte Carlo integration, nested-sampling, where a likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).

@@

As can be seen from Figure~\ref{fig:epsilon}, there is seemingly a
sweet spot for the signal error; with a $5$\% error giving an
inaccurate estimation, and with a $25$\% error giving an alright
estimation but with sudden shifts away from the ideal result. The
reason why our low error estimation performed so poorly is that the
minimum difference between our values of $y_k$ is larger than our
signal error.  As such the model could not distinguish between
different reference orientations and was essentially just picking
options at random.  This minimum difference determined by our choice of
angles and sets a lower limit of how accurate our signals can be.  In
the future we hope to look at increasing the number of reference
orientations to see what the true lower bound is.

@@

As shown in Figure~\ref{fig:beta}, eventually $\beta$ begins to over
correct our estimate to the extent that we are a full radian away from
the correct result.  The exact value of $\beta$ that will optimise our
performance varies for each choice of angle, therefore we can use it
as variable for finely tuning our results to suit our choice of angles
and our signal error.  In future, we want to also investigate the
possibility of extending our prior $p(\hat{\bf n})$ so we account for not
just the change in orientation after one time step but after 2 or 3
time steps to investigate if we can investigate the auto correlation
of our dimer's motion.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}

We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern. In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity. The MSTM package is flexible in calculating the light scattering for a variety of micro-particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy. This has potential to be applied to characterise particle size and shape that are changing while in the optical trap. The authors would like to acknowledge and thank the support for this research from the funding provided by the Leverhulme Trust.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{Acknowledgement.} The authors thank the support for this research from the funding provided by the Leverhulme Trust. \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{bib} 
\bibliographystyle{ieeetr}
